---
title: "Cell Bender Functionality & Plotting"
date: 'Compiled: `r format(Sys.Date(), "%B %d, %Y")`'
output: rmarkdown::html_vignette
theme: united
df_print: kable
vignette: >
  %\VignetteIndexEntry{Cell Bender Functionality & Plotting}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
***

<style>
p.caption {
  font-size: 0.9em;
}
</style>
 
```{r setup, include=FALSE}
all_times <- list()  # store the time for each chunk
knitr::knit_hooks$set(time_it = local({
  now <- NULL
  function(before, options) {
    if (before) {
      now <<- Sys.time()
    } else {
      res <- difftime(Sys.time(), now, units = "secs")
      all_times[[options$label]] <<- res
    }
  }
}))
knitr::opts_chunk$set(
  tidy = TRUE,
  tidy.opts = list(width.cutoff = 95),
  message = FALSE,
  warning = FALSE,
  time_it = TRUE
)
```

# Cell Bender Functionality
Cell Bender is software for elimination of technical artifacts in scRNA-seq/snRNA-seq that uses deep generative model for unsupervised removal of ambient RNA and chimeric PCR artifacts.  You can find out more info about Cell Bender from the [bioRxiv Preprint](https://www.biorxiv.org/content/10.1101/791699v1), [GitHub Repo](https://github.com/broadinstitute/CellBender), and [Documentation](https://cellbender.readthedocs.io/en/latest/).  
  
Following completion of Cell Bender scCustomize contains a couple of functions that may be helpful when creating and visualizing data in Seurat.

```{r init}
library(tidyverse)
library(patchwork)
library(viridis)
library(Seurat)
library(scCustomize)
library(qs)
```
  

```{r include=FALSE}
# Read in example Cell Bender Dual Assay Object.
astrocytes_cortex <- qread("assets/astro_nuc_seq.qs")
```


## Importing Cell Bender H5 Outputs  
The output from Cell Bender is an H5 file that is styled and can be read like 10X Genomics H5 file.  However, `Seurat::Read10X_h5` assumes that the file contains no name prefix.

However,`Read10X_h5_Multi_Directory` can be used when reading in Cell Bender files. and contains additional parameters related to Cell Bender imports.

### Cell Bender Output Structure
The following is typical output directory structure for Cell Bender files with sub-directory labeled with sample name and each file also prefixed with the sample name.

```
Parent_Directory
├── sample_01
│   └── sample_01_out_cell_barcodes.csv
│   └── sample_01_out_filtered.h5
│   └── sample_01_out.h5
│   └── sample_01_out.log
│   └── sample_01_out.pdf
└── sample_02
│   └── sample_02_out_cell_barcodes.csv
│   └── sample_02_out_filtered.h5
│   └── sample_02_out.h5
│   └── sample_02_out.log
│   └── sample_02_out.pdf
```

### Read in H5 outputs  
All we have to do is adjust the parameters to account for cell bender file names and directory structure.  

* `secondary_path = ""` as the files are directly within the immediate sub-directory
* `default_10X_path = FALSE` because these are not 10X outputs.
* `h5_filename = "_out.h5"` specifies the shared aspect of file name that is not part of sample name.  (Can also specify "_out_filtered.h5" depending on which file is desired).

```{r eval=FALSE}
cell_bender_merged <- Read10X_h5_Multi_Directory(base_path = "assets/cell_bender/", secondary_path = "", default_10X_path = FALSE, h5_filename = "_out.h5", cell_bender = TRUE, merge = TRUE, sample_names = c("WT1", "WT2"), parallel = TRUE, num_cores = 2)
```

### These matrices can then be used to create Seurat objects as desired (can also be used to create other analysis objects)  

```{r eval=FALSE}
cell_bender_seurat <- CreateSeuratObject(counts = cell_bender_merged, names.field = 1, names.delim = "_")
```

## Creating Dual Assay Objects  
Sometimes it can be helpful to create object that contains both the cell ranger values and cell bender values (we'll come to why below).  scCustomize contains a helper function `Create_CellBender_Merged_Seurat` to handle object creation in one quick step.

For this function we assume that we will use the cell calling algorithm of Cell Ranger with the modified counts for Cell Bender.  

### Read in both sets of data
```{r eval=FALSE}
cell_bender_merged <- Read10X_h5_Multi_Directory(base_path = "assets/cell_bender/", secondary_path = "", default_10X_path = FALSE, h5_filename = "_out.h5", cell_bender = TRUE, merge = TRUE, sample_names = c("WT1", "WT2"), parallel = TRUE, num_cores = 2)

cell_ranger_merged <- Read10X_h5_Multi_Directory(base_path = "assets/cell_ranger/", default_10X_path = TRUE, h5_filename = "filtered_feature_bc_matrix.h5", cell_bender = F, merge = TRUE, sample_names = c("WT1", "WT2"), parallel = TRUE, num_cores = 2)
```

### Create Dual Assay Seurat Object
To run the function the user simply needs to provide the names of the two matrices and a name for assay containing the Cell Ranger counts (by default this is named "RAW").
```{r eval=FALSE}
dual_seurat <- Create_CellBender_Merged_Seurat(raw_cell_bender_matrix = cell_bender_merged, raw_counts_matrix = cell_ranger_merged, raw_assay_name = "RAW")
```

### Optional Parameters
Users can specify any additional parameters normally passed to `Seurat::CreateSeuratObject` when using this function.
```{r eval=FALSE}
dual_seurat <- Create_CellBender_Merged_Seurat(raw_cell_bender_matrix = cell_bender_merged, raw_counts_matrix = cell_ranger_merged, raw_assay_name = "RAW", min_cells = 5, min_features = 200)
```

## Dual Assay Plotting
For Cell Bender especially, but also potentially for other assays as well, it can be helpful during analysis to plot the corrected and uncorrected counts for given feature.  scCustomize contains function `FeaturePlot_DualAssay` to make easy.

Users just need to supply the names of the two assays to plot and the features.
*NOTE: Make sure both assays have been normalized before plotting.  The function will attempt to check and make sure both assays have been normalized but has not been tested in all scenarios.*

### Example Plotting
For this example I'm using unpublished single nucleus RNA dataset from mouse cortex and have subsetted the astrocytes.

#### Ambient RNA gene
First let's plot gene that represents ambient RNA as it's restricted in expression to neurons (synaptic gene).  If Cell Bender has worked well we expect that expression of this gene will be very different between the two assays.
```{r echo=FALSE, fig.height=4, fig.width=10, fig.align='center'}
FeaturePlot_DualAssay(seurat_object = astrocytes_cortex, features = "Syt1", assay1 = "RNA", assay2 = "RAW")
```

#### Non-Ambient RNA gene
Now let's plot normally astrocyte restricted gene.  If Cell Bender has worked well we expect that expression of this gene shouldn't be very different between the two assays.
```{r echo=FALSE, fig.height=4, fig.width=10, fig.align='center'}
FeaturePlot_DualAssay(seurat_object = astrocytes_cortex, features = "Gja1", assay1 = "RNA", assay2 = "RAW")
```
