---
title: "cNMF_Functionality & Plotting"
date: 'Compiled: `r format(Sys.Date(), "%B %d, %Y")`'
output: rmarkdown::html_vignette
theme: united
df_print: kable
vignette: >
  %\VignetteIndexEntry{cNMF_Functionality & Plotting}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
***

<style>
p.caption {
  font-size: 0.9em;
}
</style>
 
```{r setup, include=FALSE}
all_times <- list()  # store the time for each chunk
knitr::knit_hooks$set(time_it = local({
  now <- NULL
  function(before, options) {
    if (before) {
      now <<- Sys.time()
    } else {
      res <- difftime(Sys.time(), now, units = "secs")
      all_times[[options$label]] <<- res
    }
  }
}))
knitr::opts_chunk$set(
  tidy = TRUE,
  tidy.opts = list(width.cutoff = 95),
  message = FALSE,
  warning = FALSE,
  time_it = TRUE
)
```


## Interaction with consensus non-negative matrix factorization (cNMF)  
scCustomize contains functions to aid in the interaction with results of consensus non-negative matrix factorization (cNMF) and Seurat objects.  For more information on cNMF see [cNMF GitHub Repo](https://github.com/dylkot/cNMF) and [cNMF Publication](https://elifesciences.org/articles/43803).  

For this example we will use example outputs from using the pbmc3k object run using the same paramters as shown in [cNMF R vignette](https://htmlpreview.github.io/?https://github.com/dylkot/cNMF/blob/main/Tutorials/R_vignette.nb.html).  

```{r}
library(ggplot2)
library(dplyr)
library(tibble)
library(magrittr)
library(Seurat)
library(scCustomize)
library(patchwork)

# Load pbmc dataset
pbmc <- pbmc3k.SeuratData::pbmc3k
```

```{r include=FALSE}
# Update pbmc check
pbmc <- UpdateSeuratObject(pbmc)

pbmc <- subset(pbmc, subset = nFeature_RNA > 200 & nCount_RNA > 200)

# change cell barcodes
cells <- Cells(pbmc)

cells_mod <- gsub(pattern = "$", replacement = "-1", x = cells)

colnames(pbmc) <- cells_mod
```

```{r fig.align='center', fig.height=8, fig.width=13, message=FALSE, warning=FALSE}
# Run basic analysis (copied from pbmc3k vignette)
pbmc <- NormalizeData(pbmc) %>% 
  FindVariableFeatures() %>% 
  ScaleData() %>% 
  RunPCA() %>% 
  FindNeighbors(dims = 1:10) %>% 
  FindClusters(resolution = 0.5) %>% 
  RunUMAP(dims = 1:10)


new.cluster.ids <- c("Naive CD4 T", "CD14+ Mono", "Memory CD4 T", "B", "CD8 T", "FCGR3A+ Mono",
    "NK", "DC", "Platelet")
names(new.cluster.ids) <- levels(pbmc)
pbmc <- RenameIdents(pbmc, new.cluster.ids)
DimPlot_scCustom(pbmc, label = TRUE, pt.size = 0.5) + NoLegend()
```



## Loading the results of cNMF  
scCustomize contains the function `Read_Add_cNMF` that can take the results from cNMF and add custom dimensionality reduction to Seurat object.  The function only requires user to supply the paths to the cNMF usage and spectra files that contain the cell and feature loadings respectively.  By default this will add new reduction "cnmf" to the Seurat object.  

```{r}
pbmc <- Read_Add_cNMF(seurat_object = pbmc, usage_file = "assets/cNMF_Example_Data/example_cNMF.usages.k_7.dt_0_1.consensus.txt", spectra_file = "assets/cNMF_Example_Data/example_cNMF.gene_spectra_score.k_7.dt_0_1.txt")
```

### Optional parameters  
There are few optional parameters in `Read_Add_cNMF` for increased user control if desired.  
* `reduction_name` & `reduction_key` These can used to either change defaults ("cnmf" and "cNMF_" respectively) or if user wants to add results from multiple cNMF runs different names can be supplied for each one.   
    - For instance save second reduction with different k value.   
* `overwrite` used to overwrite old cNMF reduction with new data using the same `reduction_name`.   

```{r eval=FALSE}
# Add result from additional cNMF run as separate reduction (for instance with different k value)
pbmc <- Read_Add_cNMF(seurat_object = pbmc, usage_file = "assets/cNMF_Example_Data/example_cNMF.usages.k_15.dt_0_1.consensus.txt", spectra_file = "assets/cNMF_Example_Data/example_cNMF.gene_spectra_score.k_15.dt_0_1.txt", reduction_name = "cnmf.alt", reduction_key = "CNMFALT_")

# Overwrite current cNMF reduction with new results but keep same `reduction_name`
pbmc <- Read_Add_cNMF(seurat_object = pbmc, usage_file = "assets/cNMF_Example_Data/example_cNMF.usages.k_15.dt_0_1.consensus.txt", spectra_file = "assets/cNMF_Example_Data/example_cNMF.gene_spectra_score.k_15.dt_0_1.txt", overwrite = TRUE)
```



## cNMF Plotting  
cNMF results can be plotted in several ways using scCustomize.  

### Plotting factor loading 
The cNMF factor cell loadings can be plotted like any other reduction using `FeaturePlot_scCustom`.  Just supply the name of the factor like any other reduction.

```{r fig.align='center', fig.height=10, fig.width=13}
FeaturePlot_scCustom(seurat_object = pbmc, features = c("cNMF_1", "cNMF_2", "cNMF_3", "cNMF_4"))
```


### Plot factor correlation  
scCustomize also has plotting function `Factor_Cor_Plot` to examine relationships between factors by plotting the correlation of feature loadings for cNMF factors.  
```{r fig.align='center', fig.height=8, fig.width=13}
Factor_Cor_Plot(object = pbmc, reduction = "cnmf")
```

By default (above) `Factor_Cor_Plot` runs hierarchical clustering on the correlations and orders the factors on the plot.  But users can also adjust this to plot the factors in numerical order.
```{r fig.align='center', fig.height=8, fig.width=13}
Factor_Cor_Plot(object = pbmc, reduction = "cnmf", cluster = FALSE)
```


`Factor_Cor_Plot` also contains a number of other optional parameters to change the output plot.
```{r eval=FALSE}
# Only plot positive correlations
Factor_Cor_Plot(object = pbmc, reduction = "cnmf", positive_only = TRUE)

# Only plot only the upper diagonal
Factor_Cor_Plot(object = pbmc, reduction = "cnmf", plot_type = "upper", cluster = FALSE)

# Only plot only the lower diagonal
Factor_Cor_Plot(object = pbmc, reduction = "cnmf", plot_type = "lower", cluster = FALSE)

# Add clustering rectangles (uses `cutree`)
Factor_Cor_Plot(object = pbmc, reduction = "cnmf", cluster_rect = T, cluster_rect_num = 6, cluster_rect_col = "magenta")
```


```{r echo=FALSE, fig.align='center', fig.height=10, fig.width=13, fig.cap="***A.** Plotting only the positive correlations (negative correlations set to 0).  **B.** Plot only the upper diagonal (must set `cluster = FALSE`).  **C.** Plot only the lower diagonal (must set `cluster = FALSE`).  **D.** Add rectangles around clustered results (uses `cutree`).*"}
# Only plot positive correlations
p1 <- Factor_Cor_Plot(object = pbmc, reduction = "cnmf", positive_only = TRUE)

# Only plot only the upper diagonal
p2 <- Factor_Cor_Plot(object = pbmc, reduction = "cnmf", plot_type = "upper", cluster = FALSE)

# Only plot only the lower diagonal
p3 <- Factor_Cor_Plot(object = pbmc, reduction = "cnmf", plot_type = "lower", cluster = FALSE)

# Add clustering rectangles (uses `cutree`)
p4 <- Factor_Cor_Plot(object = pbmc, reduction = "cnmf", cluster_rect = TRUE, cluster_rect_num = 6, cluster_rect_col = "magenta")


wrap_plots(p1, p2, p3, p4, ncol = 2) + plot_annotation(tag_levels = "A")
```



## cNMF Interaction  
scCustomize also contains functions to interact with and retrive data from the cNMF reduction.  
*NOTE: These functions are S3 generics and can be used to retrieve the same information about LIGER iNMF reductions with given LIGER object, or from iNMF reduction in Seurat object.*  

The function `Top_Genes_Factor` can be used to retrieve top loading genes from cNMF factors.  Let's check what genes load most heavily on factor 2 which we can see above loads heavily on CD14+ Monocyte cluster.  
```{r}
top_gene_factor1 <- Top_Genes_Factor(object = pbmc, factor = 2, reduction = "cnmf")

top_gene_factor1
```

By default `Top_Genes_Factor` will return the top 10 loading genes but can return any number of genes by supplying value to the `num_genes` parameter.  
```{r}
top_gene_factor1 <- Top_Genes_Factor(object = pbmc, factor = 2, num_genes = 20, reduction = "cnmf")

top_gene_factor1
```

If you want to retrieve the top loading genes for all factors within given reduction you can set `factor = "all"` and the function will return data.frame with one column per factor.
```{r eval=FALSE}
top_10_factor_df <- Top_Genes_Factor(object = pbmc, factor = "all", reduction = "cnmf")

head(top_10_factor_df, 10) 
```

```{r echo=FALSE}
top_10_factor_df <- Top_Genes_Factor(object = pbmc, factor = "all", reduction = "cnmf")

head(top_10_factor_df, 10) %>%
  kableExtra::kbl(row.names = TRUE) %>%
  kableExtra::kable_styling(bootstrap_options = c("bordered", "condensed", "responsive", "striped")) 
```

