---
title: "Customized Color Palettes & Themes"
date: 'Compiled: `r format(Sys.Date(), "%B %d, %Y")`'
output: rmarkdown::html_vignette
theme: united
df_print: kable
vignette: >
  %\VignetteIndexEntry{Customized Color Palettes & Themes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
***

<style>
p.caption {
  font-size: 0.9em;
}
</style>
 
```{r setup, include=FALSE}
all_times <- list()  # store the time for each chunk
knitr::knit_hooks$set(time_it = local({
  now <- NULL
  function(before, options) {
    if (before) {
      now <<- Sys.time()
    } else {
      res <- difftime(Sys.time(), now, units = "secs")
      all_times[[options$label]] <<- res
    }
  }
}))
knitr::opts_chunk$set(
  tidy = TRUE,
  tidy.opts = list(width.cutoff = 95),
  message = FALSE,
  warning = FALSE,
  time_it = TRUE
)
```

## Choosing Color Palettes and Themes
While the default Seurat and ggplot2 plots work well they can often be enhanced by customizing color palettes and theming options used.  To simplify/streamline this process for end users scCustomize contains a number of shortcut or helper functions for quick access to color palettes and themes that are frequently used.

```{r init}
# Load Packages
library(tidyverse)
library(patchwork)
library(scCustomize)

pbmc <- pbmc3k.SeuratData::pbmc3k.final
```


```{r include=FALSE}
pbmc <- UpdateSeuratObject(pbmc)
```


## Continuous Palettes
scCustomize has a number of quick helpers for access to palettes from viridis package.  For more information on viridis palettes and what they are see [Intro to Viridis](https://cran.r-project.org/web/packages/viridis/vignettes/intro-to-viridis.html) package vignette.

* viridis_plasma_dark_high
* viridis_plasma_light_high
* viridis_magma_dark_high
* viridis_magma_light_high
* viridis_inferno_dark_high
* viridis_inferno_light_high
* viridis_dark_high
* viridis_light_high


```{r include=FALSE}
p1 <- FeaturePlot(object = pbmc, features = "CD3E", order = T)
p2 <- FeaturePlot_scCustom(seurat_object = pbmc, features = "CD3E", order = T)
p3 <- FeaturePlot_scCustom(seurat_object = pbmc, colors_use = viridis_magma_dark_high, features = "CD3E", order = T)
p4 <- FeaturePlot_scCustom(seurat_object = pbmc, colors_use = viridis_light_high, features = "CD3E", order = T)
```

```{r echo=FALSE, fig.align='center', fig.cap="*These palettes provided better contrast for continuous palettes.\n In scCustomize the default is `viridis_plasma_dark_high`*", fig.dim=c(10, 2.5)}
wrap_plots(p1, p2, p3, p4, ncol = 4) + plot_annotation(tag_levels = 'A')
```


## Large Discrete Palettes
### `DiscretePalette_scCustomize()`
The function `DiscretePalette_scCustomize` is function inspired and modified from `Seurat::DiscretePalette()`.  Under the hood most of the palettes are called using [paletteer package](https://github.com/EmilHvitfeldt/paletteer) to call palettes without adding package dependencies.  The following palettes can be specified:

* alphabet (24)
* alphabet2 (24)
* glasbey (32)
* polychrome (36)
* stepped (24)
* ditto_seq (40)
* varibow (Dynamic)

### Palettes can either be saved to global environment as variable or called directly within other functions:
```{r eval=FALSE}
# Save as variable to global environment
polychrome_pal <- DiscretePalette_scCustomize(num_colors = 36, palette = "polychrome")

# Call within another function
DimPlot(object = obj_name, cols = DiscretePalette_scCustomize(num_colors = 26, palette = "polychrome"))
```

### The "polychrome" palette is the new default for plots with <36 levels and "varibow" when more than 36 levels
```{r fig.align='center'}
DimPlot_scCustom(seurat_object = pbmc)
```


### Visualize Palettes
To visualize these or any other palettes you can use the `PalettePlot()` function.
```{r eval=FALSE}
PalettePlot(palette = DiscretePalette_scCustomize(num_colors = 26, palette = "alphabet"))

PalettePlot(palette = c("color1", "color2", ...)

```

```{r echo=FALSE, fig.height=1,fig.width=11}
PalettePlot(palette = DiscretePalette_scCustomize(num_colors = 26, palette = "alphabet")) + ggtitle("alphabet (26 colors)")
PalettePlot(palette = DiscretePalette_scCustomize(num_colors = 26, palette = "alphabet2")) + ggtitle("alphabet2 (26 colors)")
PalettePlot(palette = DiscretePalette_scCustomize(num_colors = 32, palette = "glasbey")) + ggtitle("glasbey (32 colors)")
PalettePlot(palette = DiscretePalette_scCustomize(num_colors = 36, palette = "polychrome")) + ggtitle("polychrome (36 colors)")
PalettePlot(palette = DiscretePalette_scCustomize(num_colors = 24, palette = "stepped")) + ggtitle("stepped (24 colors")
PalettePlot(palette = DiscretePalette_scCustomize(num_colors = 40, palette = "ditto_seq")) + ggtitle("ditto_seq (40 colors)")
PalettePlot(palette = DiscretePalette_scCustomize(num_colors = 50, palette = "varibow")) + ggtitle("varibow (50 color version)")
```

### Dynamic Palettes
Among the discrete palettes provide the "varibow" palette is actually a dynamic palette whose colors will adjust across the rainbow according to `num_colors` provided:
```{r echo=FALSE, fig.height=1,fig.width=11}
PalettePlot(palette = DiscretePalette_scCustomize(num_colors = 10, palette = "varibow")) + ggtitle("varibow (10 color version)")
PalettePlot(palette = DiscretePalette_scCustomize(num_colors = 20, palette = "varibow")) + ggtitle("varibow (20 color version)")
PalettePlot(palette = DiscretePalette_scCustomize(num_colors = 30, palette = "varibow")) + ggtitle("varibow (30 color version)")
PalettePlot(palette = DiscretePalette_scCustomize(num_colors = 40, palette = "varibow")) + ggtitle("varibow (40 color version)")
```

### Shuffle Palettes
Sometimes it can be helpful to use different order for the discrete palettes depending on the end plot generated (This can be especially true with "varibow" sometimes).  `DiscretePalette_scCustomize` has optional parameter that will randomly shuffle the returned palette (see `shuffle_pal` and `seed` parameter to allow for reproducible shuffling.)
```{r eval=FALSE}
PalettePlot(palette = DiscretePalette_scCustomize(num_colors = 30, palette = "varibow")) + ggtitle("varibow (normal)")
PalettePlot(palette = DiscretePalette_scCustomize(num_colors = 30, palette = "varibow"), shuffle_pal = TRUE) + ggtitle("varibow (shuffled)")
```

```{r echo=FALSE, fig.height=1,fig.width=11}
PalettePlot(palette = DiscretePalette_scCustomize(num_colors = 30, palette = "varibow")) + ggtitle("varibow (normal)")
PalettePlot(palette = DiscretePalette_scCustomize(num_colors = 30, palette = "varibow", shuffle_pal = TRUE)) + ggtitle("varibow (shuffled)")
```


## Small Discrete Palettes
Finally scCustomize contains shortcuts for some shorter discrete palettes that can be useful for plotting groups, treatments, genotypes, etc.
```{r echo=FALSE, fig.height=1,fig.width=11}
PalettePlot(palette = NavyAndOrange()) + ggtitle("NavyAndOrange(): 2 color palette.")
PalettePlot(palette = JCO_Four()) + ggtitle("JCO_Four(): 4 color palette from ggsci package.")
PalettePlot(palette = Dark2_Pal()) + ggtitle("Dark2_Pal(): 8 color Dark2 palette from RColorBrewer package")
PalettePlot(palette = ColorBlind_Pal()) + ggtitle("ColorBlind_Pal(): 8 color modified color blind safe palette")
```

## scCustomize Color Palette Defaults
To provide better visualizations using default parameters scCustomize sets a series of defaults for plots with discrete palettes using `scCustomize_Palette()` function.

* If only two identities are being plotted will use: `NavyAndOrange()`
* If greater than 2 identities but less than or equal to 36 will use: `DiscretePalette_scCustomize(num_colors = 36, palette = "polychrome")`
* If greater than 36 identities will use dynamic varibow palette: `DiscretePalette_scCustomize(num_colors = 50, palette = "varibow", shuffle_pal = TRUE)`

```{r echo=FALSE, fig.height=1,fig.width=11}
PalettePlot(palette = NavyAndOrange()) + ggtitle("NavyAndOrange() (2 colors)")
PalettePlot(palette = DiscretePalette_scCustomize(num_colors = 36, palette = "polychrome")) + ggtitle("polychrome (36 colors)")
PalettePlot(palette = DiscretePalette_scCustomize(num_colors = 50, palette = "varibow", shuffle_pal = TRUE)) + ggtitle("varibow (50 color version)")
```


## Plotting Themes & Helpers
ggplot2 themeing works absolutely great but can also be simplified by wrapping commonly used post-plot themes into less verbose functions.
  
scCustomize contains a few customized themes and ggplot2 helpers that simplify post plot themeing/alterations.  
*NOTE: These functions can be used with any ggplot2 plot and are not limited to modifying scCustomize plots.*  

* `UnRotate_X()` Undo x-axis label rotation
* `Move_Legend()` Select new location to move legend
* `Blank_Theme()` Removes all axis labels and axes from plot
* `theme_ggprism_mod()` Modified version of `ggprism::theme_prism()`


## Suggestions/Enhancements?
If you have any ideas for additional color palettes that would be helpful to add or ggplot2 themeing functions please post as [Pull Request](https://github.com/samuel-marsh/scCustomize/pulls) on GitHub!
