---
title: "Object Conversion"
date: 'Compiled: `r format(Sys.Date(), "%B %d, %Y")`'
output: rmarkdown::html_vignette
theme: united
df_print: kable
vignette: >
  %\VignetteIndexEntry{Object Conversion}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
***

<style>
p.caption {
  font-size: 0.9em;
}
</style>
 
```{r setup, include=FALSE}
all_times <- list()  # store the time for each chunk
knitr::knit_hooks$set(time_it = local({
  now <- NULL
  function(before, options) {
    if (before) {
      now <<- Sys.time()
    } else {
      res <- difftime(Sys.time(), now, units = "secs")
      all_times[[options$label]] <<- res
    }
  }
}))
knitr::opts_chunk$set(
  tidy = TRUE,
  tidy.opts = list(width.cutoff = 95),
  message = FALSE,
  warning = FALSE,
  time_it = TRUE
)
```

# Converting between scRNA-seq package formats  
While many packages are have some object converters they are not always as feature complete as desired.  scCustomize provides a few straightforward converter functions.  
For now these functions are limited to conversions to and from Seurat and LIGER.  However, other conversions functions may be added in the future.  
```{r init}
# Load Packages
library(Seurat)
library(rliger)
library(scCustomize)
library(qs)
```


```{r eval=FALSE}
# read object
pbmc <- pbmc3k.SeuratData::pbmc3k.final
```

## Convert Seurat Objects to LIGER Objects
scCustomize contains the conversion function `as.LIGER()`.  This function has a few advantages over the conversion function `rliger::seuratToLiger()`.  
- `as.liger` will transfer meta data from Seurat to LIGER by default.  
- If a dimensionality reduction is transferred `as.liger` will also store a reduction key for use with scCustomize LIGER plotting functions to correctly set plot axes.  
- If converting a list of Seurat objects and keeping Seurat variable genes then `as.liger` can either take union or intersection of genes between objects.  
- `as.liger` has additional internal checks to ensure compatibility with both Seurat V3/4 & V5 object structures.

```{r eval=FALSE}
pbmc_liger <- as.LIGER(x = pbmc)
```


## Convert LIGER objects to Seurat Objects  
The liger/rliger package already contains a function `rliger::seuratToLiger()` to convert LIGER objects to Seurat Objects.  However, during this transfer a few things have issues crop up:  
  
* All meta data except the "dataset" column from liger_object\@cell.data is lost.
* Seurat dimensionality reduction is set to tSNE regardless of method used in LIGER analysis
* Seurat assay name not specified

scCustomize contains modified version of this function named `Liger_to_Seurat()` that solves these issues with 3 extra parameters:  
  
* `keep_meta` logical. Whether to keep meta data from the \@cell.data slot in LIGER object.  Default is TRUE.
* `reduction_label` Name of dimensionality reduction technique used (e.g., tSNE, UMAP, etc).  Ensures dim names are set correctly in Seurat object.
* `seurat_assay` Name of assay to use for data in Seurat object.  Default is "RNA".

```{r eval=FALSE}
seurat_obj <- Liger_to_Seurat(liger_object = liger_object, reduction_label = "UMAP")
```

