---
title: "Plotting #5: Spatial Plotting Functions"
date: 'Compiled: `r format(Sys.Date(), "%B %d, %Y")`'
output: rmarkdown::html_vignette
theme: united
df_print: kable
vignette: >
  %\VignetteIndexEntry{Plotting #5: Spatial Plotting Functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
***

<style>
p.caption {
  font-size: 0.9em;
}
</style>
 
```{r setup, include=FALSE}
all_times <- list()  # store the time for each chunk
knitr::knit_hooks$set(time_it = local({
  now <- NULL
  function(before, options) {
    if (before) {
      now <<- Sys.time()
    } else {
      res <- difftime(Sys.time(), now, units = "secs")
      all_times[[options$label]] <<- res
    }
  }
}))
knitr::opts_chunk$set(
  tidy = TRUE,
  tidy.opts = list(width.cutoff = 95),
  message = FALSE,
  warning = FALSE,
  time_it = TRUE
)
```


# Customizing Plots for Enhanced/Simplified Visualization in Spatial Analyses
While the default plots from Seurat and other packages are often very good they are often modified from their original outputs after plotting.  scCustomize seeks to simplify this process and enhance some of the default visualizations.  

For this tutorial, I will be utilizing two spatial datasets to illustrate scCustomize functions (both are part of [SeuratData package](https://github.com/satijalab/seurat-data)).

```{r init}
library(ggplot2)
library(dplyr)
library(magrittr)
library(patchwork)
library(viridis)
library(Seurat)
library(scCustomize)
library(qs)

# Load stxBrain Visium dataset and Slide-seq V2 ssHippo dataset
mouse_ctx <- SeuratData::LoadData("stxBrain", type = "anterior1")

slide.seq <- SeuratData::LoadData("ssHippo")
```


### In order to be compatible with this vignette we will follow Seurat vignette processing of this data.  
```{r eval=FALSE}
# Visium
mouse_ctx <- SCTransform(mouse_ctx, assay = "Spatial", verbose = FALSE)
mouse_ctx <- RunPCA(mouse_ctx, assay = "SCT", verbose = FALSE)
mouse_ctx <- FindNeighbors(mouse_ctx, reduction = "pca", dims = 1:30)
mouse_ctx <- FindClusters(mouse_ctx, verbose = FALSE)
mouse_ctx <- RunUMAP(mouse_ctx, reduction = "pca", dims = 1:30)

# Slide-seq
slide.seq <- SCTransform(slide.seq, assay = "Spatial", ncells = 3000, verbose = FALSE)
slide.seq <- RunPCA(slide.seq)
slide.seq <- RunUMAP(slide.seq, dims = 1:30)
slide.seq <- FindNeighbors(slide.seq, dims = 1:30)
slide.seq <- FindClusters(slide.seq, resolution = 0.3, verbose = FALSE)
```

```{r include=FALSE}
# Visium
mouse_ctx <- SCTransform(mouse_ctx, assay = "Spatial", verbose = FALSE)
mouse_ctx <- RunPCA(mouse_ctx, assay = "SCT", verbose = FALSE)
mouse_ctx <- FindNeighbors(mouse_ctx, reduction = "pca", dims = 1:30)
mouse_ctx <- FindClusters(mouse_ctx, verbose = FALSE)
mouse_ctx <- RunUMAP(mouse_ctx, reduction = "pca", dims = 1:30)

# Slide-seq
# NOT SURE WHY I HAVE TO SET FUTURE FOR NON-PARALLEL USE???
options(future.globals.maxSize = 1000 * 1024^2)
slide.seq <- SCTransform(slide.seq, assay = "Spatial", ncells = 3000, verbose = FALSE)
slide.seq <- RunPCA(slide.seq)
slide.seq <- RunUMAP(slide.seq, dims = 1:30)
slide.seq <- FindNeighbors(slide.seq, dims = 1:30)
slide.seq <- FindClusters(slide.seq, resolution = 0.3, verbose = FALSE)
```


## Spatial DimPlots  
Currently, spatial support in scCustomize is limited to one plotting function `SpatialDimPlot`.  scCustomize provides a matching scCustomize version of the function: `SpatialDimPlot_scCustom` to align these plots with scCustomize style and colors.  
*NOTE: If you are interested in further spatial functionality within scCustomize please post issue to GitHub or directly submit PR, as future implementations will be based on user demand.*

The issue with the default `SpatialDimPlot` are similar to those affecting the regular `DimPlot`.  Namely that the default color scheme can make it very difficult to tell spot colors from each other.  `SpatialDimPlot_scCustom()` solves these issues.


```{r echo=FALSE, fig.height=4, fig.width=13, fig.align='center', fig.cap="***A.** Default plotting results in colors that are very similar between clusters making it hard to differentiate between them on the image.  **B.** Default plot using scCustomize `SpatialDimPlot_scCustom` solves this issues matching color scheme to scCustomize package defaults.*"}
p1 <- SpatialDimPlot(object = mouse_ctx)
p2 <- SpatialDimPlot_scCustom(seurat_object = mouse_ctx)
wrap_plots(p1, p2, ncol = 2) + plot_annotation(tag_levels = "A")
```

```{r echo=FALSE, fig.height=4, fig.width=13, fig.align='center'}
p1 <- SpatialDimPlot(object = slide.seq, stroke = 0)
p2 <- SpatialDimPlot_scCustom(seurat_object = slide.seq, stroke = 0)
wrap_plots(p1, p2, ncol = 2) + plot_annotation(tag_levels = "A")
```
