---
title: "Statistics Functions"
date: 'Compiled: `r format(Sys.Date(), "%B %d, %Y")`'
output: rmarkdown::html_vignette
theme: united
vignette: >
  %\VignetteIndexEntry{Statistics Functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
***

<style>
p.caption {
  font-size: 0.9em;
}
</style>
 
```{r setup, include=FALSE}
all_times <- list()  # store the time for each chunk
knitr::knit_hooks$set(time_it = local({
  now <- NULL
  function(before, options) {
    if (before) {
      now <<- Sys.time()
    } else {
      res <- difftime(Sys.time(), now, units = "secs")
      all_times[[options$label]] <<- res
    }
  }
}))
knitr::opts_chunk$set(
  tidy = TRUE,
  tidy.opts = list(width.cutoff = 95),
  message = FALSE,
  warning = FALSE,
  time_it = TRUE
)
```



# Statistics Functions  
scCustomize contains a couple simple helper functions to return metrics of interest.

For this vignette I will be utilizing pbmc3k dataset from the SeuratData package.
```{r init}
# Load Packages
library(ggplot2)
library(dplyr)
library(magrittr)
library(Seurat)
library(scCustomize)
library(qs)

# Load example dataset for tutorial
pbmc <- pbmc3k.SeuratData::pbmc3k.final
```

```{r include=FALSE}
# Update pbmc check
pbmc <- UpdateSeuratObject(pbmc)
```

Now let's add some extra meta data for use with tutorial
```{r}
# Add mito and ribo data
pbmc <- Add_Mito_Ribo(object = pbmc, species = "human")

# Add random sample and group variables
pbmc$orig.ident <- sample(c("sample1", "sample2", "sample3", "sample4"), size = ncol(pbmc), replace = TRUE)
pbmc@meta.data$group[pbmc@meta.data$orig.ident == "sample1" | pbmc@meta.data$orig.ident == "sample3"] <- "Group 1"
pbmc@meta.data$group[pbmc@meta.data$orig.ident == "sample2" | pbmc@meta.data$orig.ident == "sample4"] <- "Group 2"

# Add dummy module score 
pbmc <- AddModuleScore(object = pbmc, features = list(c("CD3E", "CD4", "THY1", "TCRA")), name = "module_score")
```


## Cells Per Identity
It can be really helpful to know the number and percentage of cells per identity/cluster during analysis.  
scCustomize provides the `Cluster_Stats_All_Samples` function to make this easy.

```{r eval=FALSE}
cluster_stats <- Cluster_Stats_All_Samples(seurat_object = pbmc)

cluster_stats
```

```{r echo=FALSE}
cluster_stats <- Cluster_Stats_All_Samples(seurat_object = pbmc)

cluster_stats %>%
  kableExtra::kbl() %>%
  kableExtra::kable_styling(bootstrap_options = c("bordered", "condensed", "responsive", "striped")) 
```

By default `Cluster_Stats_All_Samples` uses "orig.ident" as the group variable but user can specify any `@meta.data` slot variable.
```{r eval=FALSE}
cluster_stats <- Cluster_Stats_All_Samples(seurat_object = pbmc, group_by_var = "group")

cluster_stats
```

```{r echo=FALSE}
cluster_stats <- Cluster_Stats_All_Samples(seurat_object = pbmc, group_by_var = "group")

cluster_stats %>%
  kableExtra::kbl() %>%
  kableExtra::kable_styling(bootstrap_options = c("bordered", "condensed", "responsive", "striped"))  
```


## Percent of Cells Expressing Feature(s)
It can also be informative to understand the percent of cells/nuclei that express a given feature or set of features.  
scCustomize provides the `Percent_Expressing` function to return these results.
```{r eval=FALSE}
percent_express <- Percent_Expressing(seurat_object = pbmc, features = c("CD4", "CD8A"))
```

```{r echo=FALSE}
percent_express <- Percent_Expressing(seurat_object = pbmc, features = c("CD4", "CD8A"))

percent_express %>%
  kableExtra::kbl() %>%
  kableExtra::kable_styling(bootstrap_options = c("bordered", "condensed", "responsive", "striped"))  
```


### Change grouping variable
By default the function groups expression across `@active.ident` (see above) but user can specify difference variable if desired.
```{r eval=FALSE}
percent_express <- Percent_Expressing(seurat_object = pbmc, features = c("CD4", "CD8A"), group_by = "orig.ident")
```

```{r echo=FALSE}
percent_express <- Percent_Expressing(seurat_object = pbmc, features = c("CD4", "CD8A"), group_by = "orig.ident")

percent_express %>%
  kableExtra::kbl() %>%
  kableExtra::kable_styling(bootstrap_options = c("bordered", "condensed", "responsive", "striped"))  
```

### Split within groups
User can also supply a `split_by` variable to quantify expression within group split by meta data variable
```{r eval=FALSE}
percent_express <- Percent_Expressing(seurat_object = pbmc, features = c("CD4", "CD8A"), split_by = "group")
```

```{r echo=FALSE}
percent_express <- Percent_Expressing(seurat_object = pbmc, features = c("CD4", "CD8A"), split_by = "group")

percent_express %>%
  kableExtra::kbl() %>%
  kableExtra::kable_styling(bootstrap_options = c("bordered", "condensed", "responsive", "striped"))  
```


### Set threshold of expression
Users can also supply a threshold value to quantify percent of cells expressing feature(s) above a certain threshold (be sure to note which slot is being used to quantify expression when setting thresholds).  
*NOTE: `Percent_Expressing` currently only supports single threshold across all features*
```{r eval=FALSE}
percent_express <- Percent_Expressing(seurat_object = pbmc, features = c("CD4", "CD8A"), threshold = 2)
```

```{r echo=FALSE}
percent_express <- Percent_Expressing(seurat_object = pbmc, features = c("CD4", "CD8A"), threshold = 2)

percent_express %>%
  kableExtra::kbl() %>%
  kableExtra::kable_styling(bootstrap_options = c("bordered", "condensed", "responsive", "striped"))  
```

## Calculate Summary Median Values
scCustomize contains function `Median_Stats` to quickly calculate the medians for basic QC stats (Genes/, UMIs/, % Mito/Cell).
  
### Basic Use
By default `Median_Stats` will calculate medians for the following meta data columns (if present): "nCount_RNA", "nFeature_RNA", "percent_mito", "percent_ribo", "percent_mito_ribo".

```{r eval=FALSE}
median_stats <- Median_Stats(seurat_object = pbmc, group_by_var = "orig.ident")
```

```{r echo=FALSE}
median_stats <- Median_Stats(seurat_object = pbmc, group_by_var = "orig.ident")

median_stats %>%
  kableExtra::kbl() %>%
  kableExtra::kable_styling(bootstrap_options = c("bordered", "condensed", "responsive", "striped"))  
```


### Additional Variables
In addition to default variables, users can supply their own additional meta data columns to calculate medians using the `median_var` parameter.  
*NOTE: The meta data column must be in numeric format (e.g., integer, numeric, dbl)*

```{r eval=FALSE}
median_stats <- Median_Stats(seurat_object = pbmc, group_by_var = "orig.ident", median_var = "module_score1")
```

```{r echo=FALSE}
median_stats <- Median_Stats(seurat_object = pbmc, group_by_var = "orig.ident", median_var = "module_score1")

median_stats %>%
  kableExtra::kbl() %>%
  kableExtra::kable_styling(bootstrap_options = c("bordered", "condensed", "responsive", "striped"))  
```

### Calculate Median Absolute Deviations
In addition to calculating the median values scCustomize includes function to calculate the median absolute deviation for each of those features with function `MAD_Stats`.  By setting the parameter `mad_num` the function will retrun the MAD*mad_num.

```{r eval=FALSE}
mad <- MAD_Stats(seurat_object = pbmc, group_by_var = "orig.ident", mad_num = 2)
```

```{r echo=FALSE}
mad <- MAD_Stats(seurat_object = pbmc, group_by_var = "orig.ident", mad_num = 2)

mad %>%
  kableExtra::kbl() %>%
  kableExtra::kable_styling(bootstrap_options = c("bordered", "condensed", "responsive", "striped"))  
```


## Plotting Median Data
scCustomize also contains series of functions for plotting the results of these median calculations:
  
* `Plot_Median_Genes()`
* `Plot_Median_UMIs()`
* `Plot_Median_Mito()`
* `Plot_Median_Other()`

```{r eval=FALSE}
Plot_Median_Genes(seurat_object = pbmc)

Plot_Median_Genes(seurat_object = pbmc, group_by = "group")

Plot_Median_Other(seurat_object = pbmc, median_var = "module_score1", group_by = "group")
```

```{r echo=FALSE, fig.height=7, fig.width=13, fig.align='center'}
p1 <- Plot_Median_Genes(seurat_object = pbmc)

p2 <- Plot_Median_Genes(seurat_object = pbmc, group_by = "group")

p3 <- Plot_Median_Other(seurat_object = pbmc, median_var = "module_score1", group_by = "group")

patchwork::wrap_plots(p1, p2, p3, ncol = 3)
```

